Imprecise Datapath Design: An Overclocking Approach

### 2013-05-23 TRETS2013/TRETS_overclocking.tex
Conclusion, Modified Format of Figures

Typically the maximum frequency of a RCA is determined by the longest carry propagation. Consequently, modern FPGAs offer built-in architectures for very fast ripple carry addition. For instance, the Altera Cyclone series uses fast tables~\cite{AlteraCyclone} while the Xilinx Virtex series employs dedicated multiplexers and encoders for the fast carry logic~\cite{Virtex6}. Figure~\ref{FPGA adder} illustrates the structure of an $n$-bit RCA, which is composed of $n$ serial-connected full adders (FAs) and utilizes the internal fast carry logic of the Virtex-6 FPGA.

While the fast carry logic reduces the time of each individual carry-propagation delay, the overall delay of carry-propagation will eventually overwhelm the delay of sum generation of each LUT with increasing operand word-lengths. For our initial analysis, we assume that the carry propagation delay of each FA is a constant value $\mu$, which is a combination of logic delay and routing delay, and hence the critical path delay of the RCA is $\mu_{RCA}=n\mu$, as shown in~Figure~\ref{FPGA adder}. For an $n$-bit RCA, it follows that if the sampling period $T_S$ is greater than $\mu_{RCA}$, correct results will be sampled. If, however, $T_S<\mu_{RCA}$, intermediate results will be sampled, potentially generating errors.

@@

If the input signal of a circuit is $k$ bits, truncation error occurs when the input signal is truncated from $k$ bits to $n$ bits. Under this premise, the mean value of the truncated bits at signal input ($E_{Tin}$) is given by (\ref{ET at single input}).

\begin{eqnarray}\label{ET at single input}
%\small
%\scriptsize
  \begin{split}
    E_{Tin} &= \frac{1}{2}\sum_{i=n+1}^{k}2^{-i}\\
    &= 2^{-n-1}-2^{-k-1}
%  \vspace{-1ex}
  \end{split}
\end{eqnarray}

Since we assume there are two mutually independent inputs to the RCA, the overall expectation of truncation error for the RCA is given by (\ref{TruncationError}).

\begin{eqnarray}\label{TruncationError}
%\small
%\footnotesize
  E_T=\left\{
    \begin{matrix}
        2^{-n}-2^{-k}, & \textrm{if $n<k$}\\
        0, & \textrm{otherwise}
    \end{matrix}
  \right.
\end{eqnarray}
%\normalsize

\subsection{Probabilistic Model of Overclocking Error}
\subsubsection{Generation of Overclocking Error}\label{subsub:Generation of Overclocking Error}
For a given $T_S$, the maximum length of error-free carry propagation is described by~(\ref{Max.length of carry chain}), where $f_S$ denotes the sampling frequency.

\begin{eqnarray}\label{Max.length of carry chain}
    b := \left\lceil \frac{T_S}{\mu} \right\rceil= \left\lceil \frac{1}{\mu\cdot f_S}\right\rceil
\end{eqnarray}

However, since the length of an actual carry chain during execution is dependent upon input patterns, in general, the worst case may occur rarely. To determine when this timing constraint is not met and the size of the error in this case, we expand standard results~\cite{DigitalICDesign} to the following statements, which examine carry generation, propagation and annihilation, as well as the corresponding summation results of a single bit $i$, according to the relationship between its input patterns $A_i$ and $B_i$:
%
 \begin{itemize}
   \item If $\!A_i\!=\!B_i\!=1$, a new carry chain is generated at bit $i$, and $S_i\!=\!C_{i-1}$;
   \item If $A_i\neq B_i$, the carry propagates for this carry chain at bit $i$, and $S_i\!=\!0$;
   \item If $\!A_i\!=\!B_i$, the current carry chain annihilates at bit $i$, and $S_i\!\!=\!\!1$.
 \end{itemize}

 \subsubsection{Absolute Value of Overclocking Error}
For an $n$-bit RCA, let $C_{tm}$ denote the carry chain generated at bit $S_t$ with the length of $m$ bits. For a certain $f_S$, the maximum length of error-free carry propagation, $b$, is determined through (\ref{Max.length of carry chain}). The presence of overclocking error requires $m>b$. Since the length of carry chain cannot be greater than $n$, parameters $t$ and $m$ are bounded by (\ref{t_RCA}) and (\ref{m_RCA}):

\begin{align}
%\small
%\footnotesize
%\scriptsize
  \label{t_RCA}  0&\leq t \leq n-b\\
    b&<m\leq n+1-t      \label{m_RCA}
%    \vspace{-1ex}
\end{align}

For $C_{tm}$, correct results will be generated from bit $S_t$ to bit $S_{t+b-1}$. Hence the absolute value of error seen at the output, normalized to the MSB ($2^n$), is given by (\ref{Output Error}), where $\hat{S_i}$ and $S_i$ denote the actual and error-free output of bit $i$ respectively.

\begin{eqnarray}\label{Output Error}
%\small
%\footnotesize
    e_{tm}=\frac{\left|\sum_{i=t+b}^{n}(S_i-\hat{S_i})\cdot 2^i\right|}{2^n}
%    \vspace{-1ex}
\end{eqnarray}

$S_i$ and $\hat{S_i}$ can be determined using the equations from the previous statements in Section~\ref{subsub:Generation of Overclocking Error}. In the error-free case, the carry will propagate from bit $S_t$ to bit $S_{t+m-1}$, and we will obtain $S_{t+b}=S_{t+b+1}=\cdots=S_{t+m-2}=0$ for carry propagation, and $S_{t+m-1}=1$ for carry annihilation. However, when a timing violation occurs, the carry will not propagate through all these bits. Substituting these values into (\ref{Output Error}) yields (\ref{etm}). Interestingly, the value of overclocking error has no dependence on the length of carry chain $m$.

\begin{eqnarray}\label{etm}
%\small
%\footnotesize
  \begin{split}
    e_{tm} &= \frac{\left|2^{t+m-1}-2^{t+m-2}-\dots-2^{t+b}\right|}{2^n}\\
    &= 2^{t+b-n}
  \end{split}
\end{eqnarray}

\subsubsection{Probability of Overclocking Error}
The carry chain $C_{tm}$ occurs when there is a carry generated at bit $t$, a carry annihilated at bit $t+m-1$ and the carry propagates in between. Consequently, its probability $P_{tm}$ is given by (\ref{Ptm_RCA}).

\begin{eqnarray}\label{Ptm_RCA}
%\small
%\footnotesize
  P_{tm}=P_{(A_t=B_t=1)}P_{(A_{t+m-1}=B_{t+m-1})}\cdot \prod_{i=t+1}^{t+m-2}P_{(A_i\neq B_i)}
%  \vspace{-1ex}
\end{eqnarray}

Under the assumption that $A$ and $B$ are mutually independent and uniformly distributed, we have $P_{(A_i=B_i=1)}=1/4$, $P_{(A_i\neq B_i)}=1/2$ and $P_{(A_i=B_i)}=1/2$, so $P_{tm}$ can be obtained by~(\ref{Ptm2}). Note that~(\ref{Ptm2}) takes into account the carry annihilation always occurs when $t+m-1=n$.
%\vspace{-1ex}
%
\begin{eqnarray}\label{Ptm2}
%\small
%\footnotesize
    P_{tm}=\left\{\begin{array}{ll}
      (1/2)^{m+1} & \textrm{if $t+m-1<n$}\\
      (1/2)^{m} & \textrm{if $t+m-1=n$}
    \end{array} \right.
%    \vspace{-1ex}
\end{eqnarray}

\subsubsection{Expectation of Overclocking Error}
Expectation of overclocking error can be expressed by (\ref{Eo_exp}).
\begin{eqnarray}\label{Eo_exp}
%\small
%\footnotesize
%\scriptsize
    %E_O=\sum_{t=0}^{n-b}\sum_{m=b+1}^{n-t+1}P_{tm}\cdot e_{tm}
    E_O=\sum_{t}\sum_{m}P_{tm}\cdot e_{tm}
%    \vspace{-1ex}
\end{eqnarray}

Using $P_{tm}$ and $e_{tm}$ from (\ref{etm}) and (\ref{Ptm2}) respectively, $E_O$ can be obtained by (\ref{Eo}).
%\vspace{-1ex}
\begin{eqnarray}\label{Eo}
%\small
%\footnotesize
      E_O=\left\{
        \begin{matrix}
            2^{-b}-2^{-n-1}, & \textrm{if $b\leq n$}\\
            0, & \textrm{otherwise}
        \end{matrix}
      \right.
%      \vspace{-1ex}
\end{eqnarray}

@@

\section{Carry Select Adder}\label{Section_CSA}
\subsection{Introduction}
Since the delay of RCA is determined by the length of carry chain, alternative adder architectures, such as carry select adder (CSA) have been proposed to boost performance. For CSA, the carry chain is divided into multiple overlapped sections to increase the operating speed, as shown in Fig.~\ref{Fig_CSA_Structure}. There are multiple stages within a CSA. Each stage contains two RCAs and two multiplexers, as seen in Fig.~\ref{Fig_CSA_SingleStage}. For a given input, two additions are performed simultaneously where the carry input is zero and one respectively. One of these two results is then selected according to the actual carry input. Although this structure brings performance benefits, it costs extra hardware resources compared to a standard RCA because the carry chain is duplicated. Furthermore, in FPGA technology, multiplexers are expensive. Due to this reason, we explore the trade-offs between silicon area, accuracy and performance of RCA and CSA in this section.

\begin{figure}[htbp]
    \begin{minipage}[b]{.49\textwidth}
        \centering
        \subfigure[The structure diagram of a CSA with $s$ stages.]{
            \includegraphics[width=0.95\textwidth]{./Figures/CSA_structure.pdf}
            \label{Fig_CSA_Structure}
        }%\vspace{-2ex}
        \subfigure[The structure of the $i^{th}$ stage in the CSA.]{
        %\centering
            \includegraphics[width=2.4in]{./Figures/CSA_SingleStage.pdf}
            \label{Fig_CSA_SingleStage}
        }
    \end{minipage}
    \centering
    \caption{The structure diagram of CSA.}
\end{figure}

@@

In our analysis, the stage delay of stage $i$ in the CSA refers to the combination of the $i$-bit carry propagation delay and the delay of multiplexing the carry output. For a CSA with $s$ stages~($s\geqslant 2$), let the stage delay be denoted by $d_{s-1},~\dots,~d_{0}$, where $d_{s-1}$ and $d_{0}$ represent the delay of the most significant and the least significant stages, respectively. We still follow the aforementioned assumption that the critical path delay of the CSA is due to carry propagation and multiplexing the carry output, instead of generating the sum outputs. It should be noted that unlike other stages, the least significant stage of the CSA is only built by one RCA without multiplexers, since it is directly driven by the carry input. Hence we can obtain the delay of the $i^{th}$ stage as presented in~(\ref{CSA_SingleStageDelay}), where $\mu_{c}$, $\mu_{mux}$ and $n_i$ denote the delay of $1$-bit carry propagation, the delay of multiplexing and the word-length of the $i^{th}$ stage of the CSA, respectively.
\begin{eqnarray}\label{CSA_SingleStageDelay}
  d_i=\left\{
    \begin{matrix}
      n_i\cdot \mu_{c}+(s-i)\cdot\mu_{mux}, &\textrm{if $i\in\left[1,s-1\right]$}\\
      n_0\cdot \mu_c + (s-1)\cdot\mu_{mux}, & \textrm{if $i=0$}
    \end{matrix}
  \right.
\end{eqnarray}

Under the timing-driven design environment, the delay of each stage of CSA is equalized in order to achieve the fastest operation, as presented in~(\ref{CSA_DelayEqual}).
\begin{eqnarray}\label{CSA_DelayEqual}
  d_{s-1}=d_{s-2}=\cdots=d_{0}
\end{eqnarray}

In this case, combining~(\ref{CSA_SingleStageDelay}) and (\ref{CSA_DelayEqual}) yields $n_i$, which is represented by the word-length of the most significant stage $n_{s-1}$, in (\ref{CSA_DelayRep}).

\begin{eqnarray}\label{CSA_DelayRep}
 n_i=\left\{
	\begin{matrix}
	  n_{s-1}-(s-1-i)\cdot\frac{\mu_{mux}}{\mu_c}, & \textrm{if $i\in\left[1,s-1\right]$}\\
	  n_{s-1}-(s-2)\cdot\frac{\mu_{mux}}{\mu_c}, &\textrm{if $i=0$}
	\end{matrix}
    \right.
\end{eqnarray}

Sum up $n_i$ to give the total word-length of the CSA in (\ref{CSA_WL}).
\begin{eqnarray}\label{CSA_WL}
  \begin{split}
    n_{CSA} &=\sum_{i=0}^{s-1}n_{i} \\
    &= s\cdot n_{s-1}-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+1)(s-2)}{2}
  \end{split}
\end{eqnarray}

Conventionally under a given frequency constraint, the word-length of RCA is truncated by using $n_{RCA}=b-1$ bits, where $b$ is determined by~(\ref{Max.length of carry chain}). Similarly for CSA, the word-length of each stage should be selected in order to satisfy~(\ref{TimingConstraint_CSA}).
\begin{eqnarray}\label{TimingConstraint_CSA}
  \begin{matrix}
    \forall i\in[0,s-1] &, & d_i\leqslant\frac{1}{f_s}
  \end{matrix}
\end{eqnarray}

Hence for the same frequency requirement, we can form the relationship between the delay of the most significant stage of CSA and the delay of RCA, as given by~(\ref{WL_MSB_CSA}),
\begin{eqnarray}\label{WL_MSB_CSA}
    \mu_c\cdot (b-1) = n_{s-1}\cdot\mu_c+\mu_{mux}
\end{eqnarray}

Substitute~(\ref{WL_MSB_CSA}) into (\ref{CSA_WL}) to replace $n_{s-1}$, we derive the representation of the word-length of CSA in terms of $b$, as presented in~(\ref{CSA_Timing}).
\begin{eqnarray}\label{CSA_Timing}
  n_{CSA}=s\cdot (b-1)-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+2)(s-1)}{2}
\end{eqnarray}

@@

\begin{figure}[tbp]
    \begin{minipage}[b]{0.48\textwidth}
        \centering
        \subfigure[The modeled value and the experimental results of the maximum word-length of RCA and CSA.]{
            \centering
            \includegraphics[width=0.95\textwidth]{./Figures/CSA2stage_Model_Timing.eps}
            \label{Fig_CSA2stage_Timing}
        }%
        \subfigure[Hardware resource usage for an RCA and a CSA with 2 and 3 stages.]{
            \centering
            \includegraphics[width=0.95\textwidth]{./Figures/CSA3stage_Area.eps}
            \label{Fig_CSA3stage_Area}
        }
    \end{minipage}
    \centering
    \caption{A comparison between RCA and CSA in terms of the maximum word-length of input signal and the area consumption.}
\end{figure}

@@

\subsection{Determination of the Optimum Adder Structure for Given Frequency Requirements}
First, suppose the algorithm designer wished to create a circuit that can run at a given frequency with the minimum achievable output errors and minimum resource usage, i.e. a pair of $\{Area, Frequency\}$ constraint is applied. In this case, the optimum design method is selected according to the following criteria:
\begin{itemize}
  \item Design with the minimum mean value of errors at the output is the optimum design;
  \item If multiple designs achieve the same accuracy, then the design with minimum area is the optimum design;
  \item If both the accuracy and area are identical for multiple designs, they are all treated as the optimum design.
\end{itemize}

For instance, in Fig.~\ref{Fig_CSA_LUT25} we record the mean value of error at outputs with respect to different operating frequencies for both design scenarios when the number of the available LUTs is set to 25. In this graph, we have labeled the optimum adder structure. Note that in this and all the following experiments within this section, in order to apply our models the input data is randomly generated following the uniform distribution.

\begin{figure}[th]
  \centering
  \includegraphics[width=3.3in]{./Figures/Error_LUT25.eps}
  \caption{A comparison between two design scenarios when the number of available LUT is 25. The RCA and CSA are investigated in the conventional scenario, while the RCA is explored in the proposed new scenario. The results are obtained from post place-and-route simulations on Xilinx Virtex-6 FPGAs.}
  \label{Fig_CSA_LUT25}
\end{figure}

@@

\subsection{Determination of the Optimum Adder Structure for Given Accuracy Requirements}
If the design goal is to operate the circuit as fast as possible with the minimum area whilst a certain error budget can be tolerated, the optimum design methodology can be decided as illustrated in Fig.~\ref{Fig_CSA_Tradeoff_Error}. In this situation, the error specifications are evaluated in terms of the mean relative error (MRE), as given by~(\ref{MRE}), where $E_{error}$ and $E_{out}$ refer to the mean value of error and the mean value of outputs, respectively.
\begin{eqnarray}\label{MRE}
  MRE=\left|\frac{E_{error}}{E_{out}}\right|\times 100\%
\end{eqnarray}

@@

In this CCM structure, let the two inputs of the RCA be denoted by $A_S$ and $A_O$ respectively, which are both two's complement numbers. $A_S$ denotes the ``shifted signal", with zeros padded after LSB, while $A_O$ denotes the ``original signal" with MSB sign extension. For an $n$-bit input signal, it should be noted that an $n$-bit RCA is sufficient for this operation, because no carry will be generated or propagated when adding with zeros, as shown in Fig.~\ref{CCM_fig}.
\begin{figure*}[htbp]
  \centering
  \includegraphics[width=0.99\textwidth]{./Figures/CCM_CarryChain.pdf}
  \caption{Four possible carry chain types in a constant coefficient multiplier with $n$-bit inputs. The notion $s$ denotes the shifted bits and $BP$ denotes the binary point.}
  \label{CCM_fig}
\end{figure*}

@@

\subsection{Probabilistic Model of Overclocking Error}
\subsubsection{Absolute Value of Overclocking Error}
The absolute value of overclocking error of carry chain $C_{tm}$ is increased by a factor of $2^s$ due to shifting, compared to RCA. Hence $e_{tm}$ in CCM can be modified from (\ref{etm}) to give (\ref{etm_CCM}).

\begin{eqnarray}\label{etm_CCM}
%\vspace{-1ex}
%\small
%\footnotesize
  e_{tm}=2^{t+b-n+s}
%  \vspace{-1.5ex}
\end{eqnarray}

>>>>

Typically the maximum frequency of a RCA is determined by the longest carry propagation. Consequently, modern FPGAs offer built-in architectures for very fast ripple carry addition. For instance, the Altera Cyclone series uses fast tables~\cite{AlteraCyclone} while the Xilinx Virtex series employs dedicated multiplexers and encoders for the fast carry logic~\cite{Virtex6}. Fig.~\ref{FPGA adder} illustrates the structure of an $n$-bit RCA, which is composed of $n$ serial-connected full adders (FAs) and utilizes the internal fast carry logic of the Virtex-6 FPGA.

While the fast carry logic reduces the time of each individual carry-propagation delay, the overall delay of carry-propagation will eventually overwhelm the delay of sum generation of each LUT with increasing operand word-lengths. For our initial analysis, we assume that the carry propagation delay of each FA is a constant value $\mu$, which is a combination of logic delay and routing delay, and hence the critical path delay of the RCA is $\mu_{RCA}=n\mu$, as shown in~Fig.~\ref{FPGA adder}. For an $n$-bit RCA, it follows that if the sampling period $T_S$ is greater than $\mu_{RCA}$, correct results will be sampled. If, however, $T_S<\mu_{RCA}$, intermediate results will be sampled, potentially generating errors.

@@

If the input signal of a circuit is $k$ bits, truncation error occurs when the input signal is truncated from $k$ bits to $n$ bits. Under this premise, the mean value of the truncated bits at signal input ($E_{Tin}$) is given by (\ref{ET at single input}).
%
\begin{eqnarray}\label{ET at single input}
%\small
%\scriptsize
%  \begin{split}
    E_{Tin} = \frac{1}{2}\sum_{i=n+1}^{k}2^{-i} = 2^{-n-1}-2^{-k-1}
%  \vspace{-1ex}
%  \end{split}
\end{eqnarray}

Since we assume there are two mutually independent inputs to the RCA, the overall expectation of truncation error for the RCA is given by (\ref{TruncationError}).
%
\begin{eqnarray}\label{TruncationError}
%\small
%\footnotesize
  E_T=\left\{
    \begin{matrix}
        2^{-n}-2^{-k}, & \textrm{if $n<k$}\\
        0, & \textrm{otherwise}
    \end{matrix}
  \right.
\end{eqnarray}
%\normalsize

\subsection{Probabilistic Model of Overclocking Error}
\subsubsection{Generation of Overclocking Error}\label{subsub:Generation of Overclocking Error}
For a given $T_S$, the maximum length of error-free carry propagation is described by~(\ref{Max.length of carry chain}), where $f_S$ denotes the sampling frequency.
%
\begin{eqnarray}\label{Max.length of carry chain}
    b := \left\lceil \frac{T_S}{\mu} \right\rceil= \left\lceil \frac{1}{\mu\cdot f_S}\right\rceil
\end{eqnarray}

However, since the length of an actual carry chain during execution is dependent upon input patterns, in general, the worst case may occur rarely. To determine when this timing constraint is not met and the size of the error in this case, we expand standard results~\cite{DigitalICDesign} to the following statements, which examine carry generation, propagation and annihilation, as well as the corresponding summation results of a single bit $i$, according to the relationship between its input patterns $A_i$ and $B_i$:
%
 \begin{itemize}
   \item If $A_i=B_i=1$, a new carry chain is generated at bit $i$, and $S_i=C_{i-1}$;
   \item If $A_i\neq B_i$, the carry propagates for this carry chain at bit $i$, and $S_i=0$;
   \item If $A_i=B_i$, the current carry chain annihilates at bit $i$, and $S_i=1$.
 \end{itemize}

 \subsubsection{Absolute Value of Overclocking Error}
For an $n$-bit RCA, let $C_{tm}$ denote the carry chain generated at bit $S_t$ with the length of $m$ bits. For a certain $f_S$, the maximum length of error-free carry propagation, $b$, is determined through (\ref{Max.length of carry chain}). The presence of overclocking error requires $m>b$. Since the length of carry chain cannot be greater than $n$, parameters $t$ and $m$ are bounded by (\ref{t_RCA}) and (\ref{m_RCA}):
%
\begin{align}
%\small
%\footnotesize
%\scriptsize
  \label{t_RCA}  0&\leq t \leq n-b\\
    b&<m\leq n+1-t      \label{m_RCA}
%    \vspace{-1ex}
\end{align}

For $C_{tm}$, correct results will be generated from bit $S_t$ to bit $S_{t+b-1}$. Hence the absolute value of error seen at the output, normalized to the MSB ($2^n$), is given by (\ref{Output Error}), where $\hat{S_i}$ and $S_i$ denote the actual and error-free output of bit $i$ respectively.
%
\begin{eqnarray}\label{Output Error}
%\small
%\footnotesize
    e_{tm}=\frac{\left|\sum_{i=t+b}^{n}(S_i-\hat{S_i})\cdot 2^i\right|}{2^n}
%    \vspace{-1ex}
\end{eqnarray}

$S_i$ and $\hat{S_i}$ can be determined using the equations from the previous statements in Section~\ref{subsub:Generation of Overclocking Error}. In the error-free case, the carry will propagate from bit $S_t$ to bit $S_{t+m-1}$, and we will obtain $S_{t+b}=S_{t+b+1}=\cdots=S_{t+m-2}=0$ for carry propagation, and $S_{t+m-1}=1$ for carry annihilation. However, when a timing violation occurs, the carry will not propagate through all these bits. Substituting these values into (\ref{Output Error}) yields (\ref{etm}). Interestingly, the value of overclocking error has no dependence on the length of carry chain $m$.
%
\begin{eqnarray}\label{etm}
%\small
%\footnotesize
%  \begin{split}
    e_{tm} = \frac{\left|2^{t+m-1}-2^{t+m-2}-\dots-2^{t+b}\right|}{2^n} = 2^{t+b-n}
%  \end{split}
\end{eqnarray}

\subsubsection{Probability of Overclocking Error}
The carry chain $C_{tm}$ occurs when there is a carry generated at bit $t$, a carry annihilated at bit $t+m-1$ and the carry propagates in between. Consequently, its probability $P_{tm}$ is given by (\ref{Ptm_RCA}).
%
\begin{eqnarray}\label{Ptm_RCA}
%\small
%\footnotesize
  P_{tm}=P_{(A_t=B_t=1)}P_{(A_{t+m-1}=B_{t+m-1})}\cdot \prod_{i=t+1}^{t+m-2}P_{(A_i\neq B_i)}
%  \vspace{-1ex}
\end{eqnarray}

Under the assumption that $A$ and $B$ are mutually independent and uniformly distributed, we have $P_{(A_i=B_i=1)}=1/4$, $P_{(A_i\neq B_i)}=1/2$ and $P_{(A_i=B_i)}=1/2$, so $P_{tm}$ can be obtained by~(\ref{Ptm2}). Note that~(\ref{Ptm2}) takes into account the carry annihilation always occurs when $t+m-1=n$.
%
\begin{eqnarray}\label{Ptm2}
%\small
%\footnotesize
    P_{tm}=\left\{\begin{array}{ll}
      (1/2)^{m+1} & \textrm{if $t+m-1<n$}\\
      (1/2)^{m} & \textrm{if $t+m-1=n$}
    \end{array} \right.
%    \vspace{-1ex}
\end{eqnarray}

\subsubsection{Expectation of Overclocking Error}
Expectation of overclocking error can be expressed by (\ref{Eo_exp}).
%
\begin{eqnarray}\label{Eo_exp}
%\small
%\footnotesize
%\scriptsize
    %E_O=\sum_{t=0}^{n-b}\sum_{m=b+1}^{n-t+1}P_{tm}\cdot e_{tm}
    E_O=\sum_{t}\sum_{m}P_{tm}\cdot e_{tm}
%    \vspace{-1ex}
\end{eqnarray}

Using $P_{tm}$ and $e_{tm}$ from (\ref{etm}) and (\ref{Ptm2}) respectively, $E_O$ can be obtained by (\ref{Eo}).
%
\begin{eqnarray}\label{Eo}
%\small
%\footnotesize
      E_O=\left\{
        \begin{matrix}
            2^{-b}-2^{-n-1}, & \textrm{if $b\leq n$}\\
            0, & \textrm{otherwise}
        \end{matrix}
      \right.
%      \vspace{-1ex}
\end{eqnarray}

@@

\section{Carry Select Adder}\label{Section_CSA}
\subsection{Introduction}
Since the delay of RCA is determined by the length of carry chain, alternative adder architectures, such as carry select adder (CSA) have been proposed to boost performance. For CSA, the carry chain is divided into multiple overlapped sections to increase the operating speed, as shown in Fig.~\ref{Fig_CSA_Structure}. There are multiple stages within a CSA. Each stage contains two RCAs and two multiplexers, as seen in Fig.~\ref{Fig_CSA_SingleStage}. For a given input, two additions are performed simultaneously where the carry input is zero and one respectively. One of these two results is then selected according to the actual carry input. Although this structure brings performance benefits, it costs extra hardware resources compared to a standard RCA because the carry chain is duplicated. Furthermore, in FPGA technology, multiplexers are expensive. Due to this reason, we explore the trade-offs between silicon area, accuracy and performance of RCA and CSA in this section.
\begin{figure}[htbp]
        \centering
        \subfigure[A CSA with $s$ stages.]{
            \includegraphics[width=0.6\textwidth]{./Figures/CSA_structure.pdf}
            \label{Fig_CSA_Structure}
        }%\vspace{-2ex}
        \subfigure[The $i^{th}$ stage in the CSA.]{
        \centering
            \includegraphics[width=2.2in]{./Figures/CSA_SingleStage.pdf}
            \label{Fig_CSA_SingleStage}
        }
    \caption{The structure diagram of CSA.}
\end{figure}

@@

In our analysis, the stage delay of stage $i$ in the CSA refers to the combination of the $i$-bit carry propagation delay and the delay of multiplexing the carry output. For a CSA with $s$ stages~($s\geqslant 2$), let the stage delay be denoted by $d_{s-1},~\dots,~d_{0}$, where $d_{s-1}$ and $d_{0}$ represent the delay of the most significant and the least significant stages, respectively. We still follow the aforementioned assumption that the critical path delay of the CSA is due to carry propagation and multiplexing the carry output, instead of generating the sum outputs. It should be noted that unlike other stages, the least significant stage of the CSA is only built by one RCA without multiplexers, since it is directly driven by the carry input. Hence we can obtain the delay of the $i^{th}$ stage as presented in~(\ref{CSA_SingleStageDelay}), where $\mu_{c}$, $\mu_{mux}$ and $n_i$ denote the delay of $1$-bit carry propagation, the delay of multiplexing and the word-length of the $i^{th}$ stage of the CSA, respectively.
%
\begin{eqnarray}\label{CSA_SingleStageDelay}
  d_i=\left\{
    \begin{matrix}
      n_i\cdot \mu_{c}+(s-i)\cdot\mu_{mux}, &\textrm{if $i\in\left[1,s-1\right]$}\\
      n_0\cdot \mu_c + (s-1)\cdot\mu_{mux}, & \textrm{if $i=0$}
    \end{matrix}
  \right.
\end{eqnarray}

Under the timing-driven design environment, the delay of each stage of CSA is equalized in order to achieve the fastest operation, as presented in~(\ref{CSA_DelayEqual}).
%
\begin{eqnarray}\label{CSA_DelayEqual}
  d_{s-1}=d_{s-2}=\cdots=d_{0}
\end{eqnarray}

In this case, combining~(\ref{CSA_SingleStageDelay}) and (\ref{CSA_DelayEqual}) yields $n_i$, which is represented by the word-length of the most significant stage $n_{s-1}$, in (\ref{CSA_DelayRep}).
%
\begin{eqnarray}\label{CSA_DelayRep}
 n_i=\left\{
	\begin{matrix}
	  n_{s-1}-(s-1-i)\cdot\frac{\mu_{mux}}{\mu_c}, & \textrm{if $i\in\left[1,s-1\right]$}\\
	  n_{s-1}-(s-2)\cdot\frac{\mu_{mux}}{\mu_c}, &\textrm{if $i=0$}
	\end{matrix}
    \right.
\end{eqnarray}

Sum up $n_i$ to give the total word-length of the CSA in (\ref{CSA_WL}).
%
\begin{eqnarray}\label{CSA_WL}
%  \begin{split}
    n_{CSA} =\sum_{i=0}^{s-1}n_{i}= s\cdot n_{s-1}-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+1)(s-2)}{2}
%  \end{split}
\end{eqnarray}

Conventionally under a given frequency constraint, the word-length of RCA is truncated by using $n_{RCA}=b-1$ bits, where $b$ is determined by~(\ref{Max.length of carry chain}). Similarly for CSA, the word-length of each stage should be selected in order to satisfy~(\ref{TimingConstraint_CSA}).
%
\begin{eqnarray}\label{TimingConstraint_CSA}
  \begin{matrix}
    \forall i\in[0,s-1] &, & d_i\leqslant\frac{1}{f_s}
  \end{matrix}
\end{eqnarray}

Hence for the same frequency requirement, we can form the relationship between the delay of the most significant stage of CSA and the delay of RCA, as given by~(\ref{WL_MSB_CSA}),
%
\begin{eqnarray}\label{WL_MSB_CSA}
    \mu_c\cdot (b-1) = n_{s-1}\cdot\mu_c+\mu_{mux}
\end{eqnarray}

Substitute~(\ref{WL_MSB_CSA}) into (\ref{CSA_WL}) to replace $n_{s-1}$, we derive the representation of the word-length of CSA in terms of $b$, as presented in~(\ref{CSA_Timing}).
%
\begin{eqnarray}\label{CSA_Timing}
  n_{CSA}=s\cdot (b-1)-\frac{\mu_{mux}}{\mu_{c}}\cdot\frac{(s+2)(s-1)}{2}
\end{eqnarray}

@@

\begin{figure}[htbp]
        \centering
        \subfigure[\scriptsize The modeled value and the experimental results of the maximum word-length of RCA and CSA.]{
            \includegraphics[width=3.5in]{./Figures/CSA2stage_Model_Timing.eps}
            \label{Fig_CSA2stage_Timing}
        }
        \subfigure[\scriptsize Hardware resource usage for an RCA and a CSA with 2 and 3 stages.]{
            \includegraphics[width=3.5in]{./Figures/CSA3stage_Area.eps}
            \label{Fig_CSA3stage_Area}
        }
    \caption{A comparison between RCA and CSA in terms of the maximum word-length of input signal and the area consumption.}
\end{figure}

@@

\subsection{Determination of the Optimum Adder Structure for Given Frequency Requirements}
First, suppose the algorithm designer wished to create a circuit that can run at a given frequency with the minimum achievable output errors and minimum resource usage, i.e. a pair of $\{Area, Frequency\}$ constraint is applied. In this case, the optimum design method is selected according to the following criteria:

\begin{itemize}
  \item Design with the minimum mean value of errors at the output is the optimum design;
  \item If multiple designs achieve the same accuracy, then the design with minimum area is the optimum design;
  \item If both the accuracy and area are identical for multiple designs, they are all treated as the optimum design.
\end{itemize}

For instance, in Fig.~\ref{Fig_CSA_LUT25} we record the mean value of error at outputs with respect to different operating frequencies for both design scenarios when the number of the available LUTs is set to 25. In this graph, we have labeled the optimum adder structure. Note that in this and all the following experiments within this section, in order to apply our models the input data is randomly generated following the uniform distribution.
%
\begin{figure}[t]
  \centering
  \includegraphics[width=3.3in]{./Figures/Error_LUT25.eps}
  \caption{A comparison between two design scenarios when the number of available LUT is 25. The RCA and CSA are investigated in the conventional scenario, while the RCA is explored in the proposed new scenario. The results are obtained from post place-and-route simulations on Xilinx Virtex-6 FPGAs.}
  \vspace{-.5ex}
  \label{Fig_CSA_LUT25}
\end{figure}

@@

\subsection{Determination of the Optimum Adder Structure for Given Accuracy Requirements}
If the design goal is to operate the circuit as fast as possible with the minimum area whilst a certain error budget can be tolerated, the optimum design methodology can be decided as illustrated in Fig.~\ref{Fig_CSA_Tradeoff_Error}. In this situation, the error specifications are evaluated in terms of the mean relative error (MRE), as given by~(\ref{MRE}), where $E_{error}$ and $E_{out}$ refer to the mean value of error and the mean value of outputs, respectively.
%
\begin{eqnarray}\label{MRE}
  MRE=\left|\frac{E_{error}}{E_{out}}\right|\times 100\%
\end{eqnarray}

@@

In this CCM structure, let the two inputs of the RCA be denoted by $A_S$ and $A_O$ respectively, which are both two's complement numbers. $A_S$ denotes the ``shifted signal", with zeros padded after LSB, while $A_O$ denotes the ``original signal" with MSB sign extension. For an $n$-bit input signal, it should be noted that an $n$-bit RCA is sufficient for this operation, because no carry will be generated or propagated when adding with zeros, as shown in Fig.~\ref{CCM_fig}.

\begin{figure}[htbp]
  \vspace{-2ex}
  \centering
  \includegraphics[width=\textwidth]{./Figures/CCM_CarryChain.pdf}
  \vspace{-2ex}
  \caption{Four possible carry chain types in a constant coefficient multiplier with $n$-bit inputs. The notion $s$ denotes the shifted bits and $BP$ denotes the binary point.}
  \label{CCM_fig}
\end{figure}

@@

\subsection{Probabilistic Model of Overclocking Error}
\subsubsection{Absolute Value of Overclocking Error}
The absolute value of overclocking error of carry chain $C_{tm}$ is increased by a factor of $2^s$ due to shifting, compared to RCA. Hence $e_{tm}$ in CCM can be modified from (\ref{etm}) to give (\ref{etm_CCM}).
%
\begin{eqnarray}\label{etm_CCM}
%\vspace{-1ex}
%\small
%\footnotesize
  e_{tm}=2^{t+b-n+s}
%  \vspace{-1.5ex}
\end{eqnarray}
